Model Checking

### 2023-10-05 owen/Write Ups/03_existing_models.qmd
adding indiv level permutations also

```{r}
#| eval: false

@@

One difference with this model is that we now need to consider the 
history of all opponents played by focal players also.
For the chess example where many players are involved, this leads to thousands
of opponents, many who plan only a small number of games.

@@

Only consider a single game in each window?

@@

tidy_games_mod

@@

```{r, message=FALSE}
stan_data_mod <- list(N = nrow(init_data_mod),
                  J = length(top_players),
                  y = init_data_mod$focal_result,
                  id = init_data_mod$focal_id,
                  colour = init_data_mod$focal_white,
                  elo = init_data_mod$elo_diff,
                  win_prop = init_data_mod$ave_prop)

@@

fit3_mod <- mod$sample(data = stan_data_mod,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4,
                  refresh = 100)
```

@@

This quick permutations seem to indicate that the is evidence for these 
effects, which is not present when the results are randomly permuted 
instead.

@@

for(i in seq_along(beta_var)){
  plot_curr <- perm_results %>% 
  filter(variable == beta_var[i]) %>% 
  ggplot(aes(mean)) +
  geom_histogram() +
  labs(title = paste0("Permutations of ", beta_var[i]),
       subtitle = "True data appears different",
       x = alpha_var[i]) +
  geom_vline(data = orig %>% 
             filter(variable == beta_var[i]),
             aes(xintercept = mean), col = "red")
  print(plot_curr)
}

@@

>>>>

```{r}
#| eval: true

@@

One difference with this model is that we now need to consider the 
history of all opponents played by focal players also.
For the chess example where many players are involved, this leads to thousands
of opponents, many who play only a small number of games.

@@

Only consider a single game in each window?
This would be done by only using the final game in each subsequent
window of size $n$, reducing the number of observations in 
the data by a factor of $n$ essentially. Initially it was
thought that this was being considered in the Deng et al. paper,
but that was incorrect.

@@

# tidy_games_mod

@@

```{r, message=FALSE, }
stan_data_mod <- list(N = nrow(init_data_mod),
                  J = length(top_players),
                  y = init_data_mod$focal_result,
                  id = init_data_mod$focal_id,
                  colour = init_data_mod$focal_white,
                  elo = init_data_mod$elo_diff,
                  win_prop = init_data_mod$ave_prop)

@@

fit3_mod <- mod$sample(data = stan_data_mod,
                  seed = 123,
                  chains = 4,
                  parallel_chains = 4)
```

@@

This quick permutation seem to indicate that there is evidence for these 
effects, which is not present when the results are randomly permuted 
instead.

@@

for(i in seq_along(beta_var)){
  plot_curr <- perm_results %>% 
  filter(variable == beta_var[i]) %>% 
  ggplot(aes(mean)) +
  geom_histogram() +
  labs(title = paste0("Permutations of ", beta_var[i]),
       subtitle = "True data appears different",
       x = beta_var[i]) +
  geom_vline(data = orig %>% 
             filter(variable == beta_var[i]),
             aes(xintercept = mean), col = "red")
  print(plot_curr)
}

@@

#### Checking Permuting within Players only

```{r read the data in}
perm_path <- here("owen", "cluster_scripts",
                  "perm_outputs", "perm_outputs_indiv")

perm_results <- sims %>% 
  map_dfr(~readRDS(here("owen", "cluster_scripts",
                        "perm_outputs", "perm_outputs_indiv",
                        .x))) %>% 
  select(variable, mean, median, sim_id)

```{r plot global pars}

## have orig from above

## mean winner effect

perm_results %>% 
  filter(variable == "mu2") %>% 
  ggplot(aes(mean)) +
  geom_histogram() +
  labs(title = "Permutations of Mu_2",
       subtitle = "This is the mean of the player effects",
       x = "mu2") +
  geom_vline(data = orig %>% 
               filter(variable == "mu2"),
             aes(xintercept = mean), col = "red")

## effect of playing as white

perm_results %>% 
  filter(variable == "gamma1") %>% 
  ggplot(aes(mean)) +
  geom_histogram() +
  labs(title = "Permutations of gamma_1",
       subtitle = "True data appears similar",
       x = "gamma_1") +
  geom_vline(data = orig %>% 
               filter(variable == "gamma1"),
             aes(xintercept = mean), col = "red")

## effect of elo difference

```{r check alpha first}

## some of these still significant here, need to check it

```{r check beta also}

for(i in seq_along(beta_var)){
  plot_curr <- perm_results %>% 
  filter(variable == beta_var[i]) %>% 
  ggplot(aes(mean)) +
  geom_histogram() +
  labs(title = paste0("Permutations of ", beta_var[i]),
       subtitle = "True data appears different",
       x = beta_var[i]) +
  geom_vline(data = orig %>% 
             filter(variable == beta_var[i]),
             aes(xintercept = mean), col = "red")
  print(plot_curr)
}
